---
title: "PSTAT 10: Worksheet 00"
subtitle: "Sarfehnia, Romina: rsarfehnia"
author: "Did not collaborate with anyone."
date: "2026-10-05"
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

Work in groups of at most 3 assigned by your TA. Take turns to explain each exercise to your group mates before proceeding to the next exercise as a group. 

---

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 

---

## Exercise 1: `R` as a calculator

Recall that `R` can be used as a calculator! The following arithmetic operators are understood by `R`:

- $+$ addition 
- $-$ subtraction
- $*$ multiplication
- / division

There are many built-in functions to make calculations easier. For example, calculate the $\sqrt{256}$ using the square root function `sqrt()` \

**(a)** Calculate $\displaystyle \sqrt{ \frac{3^{0.527}}{1 + \ln\left( \frac{11}{3} \right)}  }$

`r bfcolor("Solutions:", "blue")` 

```{r basicmath}
sqrt(3^0.527 / (1 + log(11/3)))
```

**(b)** Calculate $\displaystyle \frac{1}{1 + \frac{2}{1 + \frac{3}{1 + 4}}}$

`r bfcolor("Solutions:", "blue")`

```{r basicmath2}
1 / (1 + (2 / (1 + (3 / ( 1 + 4)))))
```
---

# Exercise 2: Variables and assignment

For this problem, it is very important you follow the steps in order.

**(a)** Create a new variable called `x`, and assign it the value 4.

`r bfcolor("Solutions:", "blue")`

```{r variableX}
x <- 4
```

**(b)** Now assign the value 2 to `x`.

`r bfcolor("Solutions:", "blue")`

```{r variableX2}
x <- 2
```

**(c)** Write `R` code to add 5 to whatever is the value of `x`. Don't store this new value, and don't just print the number "7".

`r bfcolor("Solutions:", "blue")`

```{r adding5}
x + 5
```

**(d)** **Before you run any code,** what do you think the result of `x / 2` will be?

`r bfcolor("Solutions:", "blue")`

I believe the result of `x / 2` will divide the current value of x divided by 2, however, I do have some experience with other programming languages such as Python and if we follow the rounding rules in Python, if the number given is not a whole number (integer), then we round DOWN ONLY IF the sign '/' is used. I will guess that the same logic is applied in R, therefore my guess would be 3.


**(e)** Run the code  `x / 2`  and comment on the result. Did you get `3.5` which is `7/2` ? Why or why not?

`r bfcolor("Solutions:", "blue")`

```{r dividing2}
x / 2
```
I did not get 3.5. This is because we did not store the variable in the previous question, so in reality, x is still equal to 2. 2/2 is equal to 1.

**(f)** **Before you run any code,** what do you think the result of `x / 0` will be?

`r bfcolor("Solutions:", "blue")`

I believe the result of `x / 0` will give us an error.

**(g)** Finally, run the code `x / 0` and comment on the result. Is this surprising?

`r bfcolor("Solutions:", "blue")`

```{r dividing0}
x / 0
```
The result of `x / 0` gave us infinity and this is surprising to me, but not too surprising, since any number divided by 0 is undefined. I guess R does not throw an error however, unlike Python.

---

# Exercise 3: `R` Datasets Package
When you start an R session, R automatically installs some packages. One such package that is loaded is the Base R `datasets` . 


**(a)** Read the help page for the `datasets` package. Note down the command you used as your solution to this exercise.

`r bfcolor("Solutions:", "blue")`

```{r datasets}
?datasets
```

**(b)** **List of datasets that come with R Datasets Package: ** Use the  help page for the `datasets` package to look at all the datasets that are contained in the `datasets` package. Note down the command you used as your solution to this exercise. Write down as a list, three packages that seem interesting to you and what makes them interesting to you.  

`r bfcolor("Solutions:", "blue")`

```{r datasets2}
?datasets
```
1. US Airport dataset (airports::usairports) seems interesting to me, since I would like to know all of the airports in the United States.
2. Car Speeding and Warning Signs (boot::amis) seems interesting to me, since I would like to know how fast drivers go and the warning signs that are on the road.
3. Data from a Simulated Motorcycle Accident (boot::motor) seems interesting to me, since I would like how often motorcyles crash and the data that is related to a motorcycle accident.

---

# Exercise 4: Tokyo Drift

In this exercise, we will play around with a relatively famous `R` dataset called `mtcars.` 

**(a)** Access the help file on `mtcars`. Do you need to install and/or load any packages to access the `mtcars` dataset? If so, install and/or load those packages.

`r bfcolor("Solutions:", "blue")`

```{r mtcars}
?mtcars
```


**(b)** In what year was this data collected? (As Data Scientists, it is very important to know as much about our data as possible *before* beginning to run analyses! Formally, in datascience, this help file about a dataset is called `Data Codebook`)

`r bfcolor("Solutions:", "blue")`

The data was collected in 1974.

**(c)** If a particular car has a `carb` value of 4, what does this mean (in words)? (In other words, what does the variable `carb` measure?) _You do not need to run any code for this exercise._

`r bfcolor("Solutions:", "blue")`

If a particular car has a 'carb' value of 4, this means that the number of carburetors the car has is 4.


**(d)** Take a look at the data with the `View` command. Do you notice any `NA` values? Notice how  the column names correspond to the ones you saw in the `Data codebook` tab earlier. 

`r bfcolor("Solutions:", "blue")`

```{r view, eval = FALSE}
View(mtcars)
```
I did not notice any 'NA' values.

**(e)** Compute the mean number of miles per gallon of the cars included in the `mtcars` dataset. Assign this mean value to a variable. Be sure to name the variable appropriately. 

`r bfcolor("Solutions:", "blue")`

```{r mean}
mean <- mean(mtcars$mpg)
mean
```

**(f)** Sort the amount of gross horsepower in decreasing order. _Use the `sort` function._

`r bfcolor("Solutions:", "blue")`

```{r sort}
sort(mtcars$hp, decreasing = TRUE)
```

**(g)** **FTC: Optional, Extra** Using the print and paste function, print "The mean number of miles per gallon of cars included in the mtcars dataset is xxxx." Your statement should have xxx replaced by the value of mean you calculated earlier. 

_Hint :_ Check the help for the paste function and search for the line that begins with _paste("1st", ..._ or _paste("The value is",...._

`r bfcolor("Solutions:", "blue")`

```{r PrintAndPaste}
print(paste("The mean number of miles per gallon of cars included in the mtcars dataset is", mean))
```

---

# Exercise 5: Learning Gains

Please remember to fill out these learning gains before showing your TA your work to get Worksheet credit!

- Understood what YAML stands for and means
- How to name and assign variables 
- How to access and use packages and datasets including: finding the mean and printing and pasting specific columns of the datasets
