---
title: "PSTAT 10: Worksheet 2"
subtitle: "Sarfehnia, Romina: rsarfehnia"
author: "Did not collaborate with anyone."
date: "2026-10-05"
output:
  html_document:
    theme: readable
    toc: yes
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = F}
bfcolor <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{\\textbf{%s}}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'><b>%s</b></span>", color, x)
  } else x
}
```

Work in groups of at most 3 assigned by your TA. Take turns to explain each exercise to your group mates before proceeding to the next exercise as a group. 

---

## Exercise 0: YAML

Update the YAML to reflect your own information (i.e. name, collaborators, etc.). If you did not collaborate with anyone, note that down in the YAML. 

---

## Exercise 1: Element-ary, Dear Watson

Create a vector `x` containing the elements 8, 14, 6, 11, 2, 1, 23, 4, 7

Write code to:

**(a)** extract the 6th element of `x`.

`r bfcolor("Solutions:", "blue")`

```{r extraction}
x <- c(8, 14, 6, 11, 2, 1, 23, 4, 7)
x[6]
```

**(b)** extract the 2nd , 3rd and 4th elements of `x`

`r bfcolor("Solutions:", "blue")`

```{r CoupleExtractions}
x[2]
x[3]
x[4]
```

**(c)** find the median of `x`

`r bfcolor("Solutions:", "blue")`

```{r median}
median(x)
```

**(d)** determine the length of `x`

`r bfcolor("Solutions:", "blue")`

```{r length}
length(x)
```

**(e)** **(Optional)** Assign the name of nine of your favorite foods as names attribute of this vector `x`. Verify that that the names have been assigned correctly.

`r bfcolor("Solutions:", "blue")`

```{r name}
names(x) <- c('Ghormeh Sabzi', 'Kime', 'Salmon', 'Avocado Sandwiches', 'Tacos', 'Burritos', 'Potstickers', 'Dumplings', 'Adasi')
x
```

\

---

## Exercise 2: Rep-and-Roll

Using `rep()` function, create the following vectors:

**(a)** 5 5 5 5 5 5 5

`r bfcolor("Solutions:", "blue")`

```{r rep}
c(rep(5, times = 7))
```

**(b)** 1 1 1 4 4 4 9 9 9 

`r bfcolor("Solutions:", "blue")`

```{r rep2}
c(rep(c(1, 4, 9), each = 3))
```

\

---

## Exercise 3: Combined Intellect

Given variables `p <- c(1,4,5,7,7)` and `q <- c(3,5,4,1,5)`. 

**(a)** Create a vector `z` by combining the vectors `p` and `q`.

`r bfcolor("Solutions:", "blue")`

```{r combination}
p <- c(1, 4, 5, 7, 7)
q <- c(3, 5, 4, 1, 5)
z <- c(p, q)
z
```

**(b)** Explain the difference in the results of `p + q` and
`sum(p + q)`

`r bfcolor("Solutions:", "blue")`

```{r AddingVectors}
p + q
sum(p + q)
```
The difference between p + q and sum(p + q) is that p + q adds up the vectors p and q at the matching indices. For example, p[1] is added to q[1] and the sum is stored in z[1]. However, sum(p + q) adds all the values within vector p and vector q, making vector z just one whole sum.

\

---

## Exercise 4: :)

Rewrite the following code using the colon operator.

```{r, eval = F}
x <- c(10, 11, 12, 13, 14, 13, 12, 11, 10)
```

`r bfcolor("Solutions:", "blue")`

```{r concatenation}
x <- c(10, 11, 12, 13, 14, 13, 12, 11, 10)
x <- c(x[1:9])
x
```

\

---

## Exercise 5: Rep-and-Roll, Electric Boogaloo

Use the replicate function `rep()` to generate the following sequence:

7 7 8 8 9

(Note: there is not an error here. The required sequence has only one instance of ‘9’)

`r bfcolor("Solutions:", "blue")`

```{r repetition}
rep(c(7, 8, 9), each = 2)
```

---

# Last Exercise: Learning Gains

Please remember to fill out these learning gains before showing your TA your work to get Worksheet credit!

- How to create a VECTOR in R in MULTIPLE different ways
- Basic functions of R such as ADDING and REPETITION of values
- How to calculate the MEDIAN of vector values and EXTRACTING values from the vectors

